### 2023-02-08 paper-thesis/chapters/02-background-spoofax.tex
Work on CoC chapter

\section[Scope graphs]{\label{sec:statix}Scope graphs\protect\footnote{Note: This section is adapted from "Composable Type System Specification using Heterogeneous Scope Graphs".~\cite[sect. 4.1.2]{Zwaan21}}}

>>>>

\section[Scope graphs]{\label{sec:statix}Scope Graphs\protect\footnote{Note: This section is adapted from "Composable Type System Specification using Heterogeneous Scope Graphs".~\cite[sect. 4.1.2]{Zwaan21}}}

### 2023-02-08 paper-thesis/chapters/10-coc.tex
Improve the variable capture chapter
Improve booleans chapter
Work on CoC chapter

\chapter{Calculus of Constructions}
\label{chap:baselang}

In this section, we will describe how to implement a dependently typed language in Statix. In section \ref{sec:coc-syntax} we will describe the syntax of the language, then in section \ref{sec:coc-scopes} we will describe how scope graphs are used to type check the language. Section \ref{sec:coc-dynsyms} describes the dynamic semantics of the language, and finally \ref{sec:coc-typecheck} how to type check the language.

\section{The language}
\label{sec:coc-syntax}

@@

\begin{figure}[h]
\begin{lstlisting}
Expr.Let = "let" ID "=" Expr ";" Expr
Expr.Type = "Type"
Expr.Var = ID
Expr.FnType = ID ":" Expr "->" Expr {right}
Expr.FnConstruct = "\\" ID ":" Expr "." Expr
Expr.FnDestruct = Expr Expr {left}
\end{lstlisting}
\caption{The concrete syntax for the base language. $\co{FnConstruct}$ is a lambda function, $\co{FnDestruct}$ is application of a lambda function.}
\label{fig:syntax}
\end{figure}

Please observe that the syntax definition does not have a separate sort for types, as types may be arbitrary expressions in a dependently typed language. Furthermore, \verb|FnType| assigns a name to its first argument to allow the return type of the function to depend on the argument type.

@@

To type check the base language, we need to store information about the names that are in scope at each point in the program. There are two different cases, names that do not have a known value (only a type), which are names created by function arguments and dependent function types, and names that do have a known value, which are names created by let bindings.\footnote{In non-dependent languages there is no such distinction, but because we may need \emph{the value} of a binding to compare types, this is needed in dependently typed languages.}

@@

\begin{lstlisting}
sPutType  : scope * ID * Expr -> scope
sPutSubst : scope * ID * (scope * Expr) -> scope
sGetName  : scope * ID -> NameEntry
sEmpty    : -> scope
\end{lstlisting}
The \verb|sPutType| and \verb|sPutSubst| predicates generate a new scope given a parent scope and a type or a substitution respectively. To query the scope graph we use \verb|sGetName|, which will return the closest \verb|NameEntry| with a matching same. Finally, \verb|sEmpty| returns a fresh empty scope.

@@

\section{Beta equality}
\label{sec:coc-betaeq}

@@

\section{Type checking the Calculus of Constructions}
\label{sec:coc-typecheck}

@@

\section{Discussion of a substitution-based approach}
\label{sec:coc-subst}

@@

		\inferrule{
			\bhr
			{\scope{s}{f}}{}
			{\scope{s_{f'}}{f'}}
			\\ \nexists x e_1 e_2. f' = \co{FnConstruct}(x, e_1, e_2)
		} {
			\bhr
			{\scope{s}{\co{FnDestruct}(f, a)}}{}
			{\scope{s}{\co{FnDestruct}(f, a)}}{}
		}

>>>>

\chapter{Calculus of Constructions in Statix}
\label{chap:baselang}

In this section, we will describe how to implement a dependently typed language in Statix. In section \ref{sec:coc-syntax} we will describe the syntax of the language, then in section \ref{sec:coc-scopes} we will describe how scope graphs are used to type check the language. Section \ref{sec:coc-dynsyms} describes the dynamic semantics of the language, and finally \ref{sec:coc-typecheck} how to type check the language. This chapter is the main contribution of this thesis.

\section{The Language}
\label{sec:coc-syntax}

@@

\begin{figure}[h]
\begin{lstlisting}
Expr.Type = "Type"
Expr.Var = ID
Expr.FnType = ID ":" Expr "->" Expr {right}
Expr.FnConstruct = "\\" ID ":" Expr "." Expr
Expr.FnDestruct = Expr Expr {left}
Expr.Let = "let" ID "=" Expr ";" Expr
\end{lstlisting}
\caption{The concrete syntax for the base language. $\co{FnConstruct}$ is a lambda function, $\co{FnDestruct}$ is application of a lambda function.}
\label{fig:syntax}
\end{figure}

There is only one sort: Expr. The syntax definition does not have a separate sort for types, as types may be arbitrary expressions in a dependently typed language. The following constructors exist:
\begin{itemize}
	\item \verb|Type| is the Type of Types.
	\item \verb|Var| is a variable, it uses the lexical sort \verb|ID|, which is defined as \verb|[a-zA-Z\_][a-zA-Z0-9\_]*|.
	\item \verb|FnType| is the type of a function. It assigns a name to its first argument to allow the return type of the function to depend on the argument type. It is right associative, meaning \verb|A -> B -> C| is interpreted as \verb|A -> (B -> C)|. 
	\item \verb|FnConstruct| creates an anonymous function (lambda function).
	\item \verb|FnDestruct| applies a function to an argument. It is left associative, meaning \verb|a b c| is interpreted as \verb|(a b) c|.
	\item \verb|Let| is a let binding. It introduces a substitutable variable.
\end{itemize}

@@

To type check the base language, we need an environment to store information about the names that are in scope at each point in the program. There are two different types of names that we may want to store, names that do not have a known value (only a type), which are names created by function arguments and dependent function types, and names that do have a known value, which are names created by let bindings.\footnote{In non-dependent languages there is no such distinction, but because we may need \emph{the value} of a binding to compare types, this is needed in dependently typed languages.}

@@

\begin{lstlisting}
sPutType  : scope * ID * Expr -> scope
sPutSubst : scope * ID * (scope * Expr) -> scope
sGetName  : scope * ID -> NameEntry
sEmpty    : -> scope
\end{lstlisting}
The \verb|sPutType| and \verb|sPutSubst| predicates generate a new scope given a parent scope and a type or a substitution respectively. These return a scope that represent an environment that has been extended with the new name. To query the scope graph we use \verb|sGetName|, which will return the closest \verb|NameEntry| with a matching name. Finally, \verb|sEmpty| returns a fresh empty scope.

@@

\section{Beta Equality}
\label{sec:coc-betaeq}

@@

\section{Type Checking}
\label{sec:coc-typecheck}

@@

\section{Discussion of a Substitution-Based Approach}
\label{sec:coc-subst}

@@

		\inferrule{
			\bhr
			{\scope{s}{f}}{}
			{\scope{s_{f'}}{f'}}
			\\ \nexists x \, e_1 \, e_2. \, \, f' = \co{FnConstruct}(x, e_1, e_2)
		} {
			\bhr
			{\scope{s}{\co{FnDestruct}(f, a)}}{}
			{\scope{s}{\co{FnDestruct}(f, a)}}{}
		}

### 2023-02-08 paper-thesis/chapters/23-adding-universes.tex
Work from yesterday
Minor writing changes
Work on CoC chapter

Our language so far has the property that the type of \verb|Type| is \verb|Type|, this is called \emph{type in type}. This allows for \emph{Girard's paradox}, which means that the logic created by this language is inconsistent, we can prove false.

@@

\section{Implementing universes in the CoC}

@@

\begin{figure}[h]
	\figuresection[\fbox{$\toe{s}{e}{t}$}]{Type checking rules with universes}
	\begin{mathpar}
		\mprset{vskip=0.7ex}
		% Type rule
		\inferrule{
		} {
			\toe{s}{\co{Type}(u)}{\co{Type}(u+1)}
		}

@@

\newpage
\section{Implementing universes with inductive data types}

>>>>

Our language so far has the property that the type of \verb|Type| is \verb|Type|, this is called \emph{type in type}. This allows for \emph{Girard's paradox}, which means that the logic created by this language is inconsistent: we can prove false.

@@

\section{Implementing Universes in the CoC}

@@

\begin{figure}[ht]
	\figuresection[\fbox{$\toe{s}{e}{t}$}]{Type checking rules with universes}
	\begin{mathpar}
		\mprset{vskip=0.7ex}
		% Type rule
		\inferrule{
		} {
			\toe{s}{\co{Type}(u)}{\co{Type}(u+1)}
		}

@@

\newpage
\section{Implementing Universes with Inductive Data Types}
